Geometron: A Universal Language for Graphical Communication

### 2018-09-09 paper1.tex
working on the workflow, recompiling repeatedly

\title{Geometron: A Language for Graphical Communication}
\author{Lafe Spietz}

@@

    The choices made here are in some sense arbitrary: 8 could just as well be 4 or 16 or even some random prime number.  But the point of this approach to language design is to choose numbers and structures which are well suited to easily being interacted with by the human mind.  A 8x8x8 cube can be divided into``tablets'', each of which is 8X in size.

@@

\section{Workflow and Applications}

@@

@@

Read this paper(or skim it and look at the pictures) 
download free graphics for use in your documents from graphics library
try out basic movements, drawing actions, make circles, lines, and gates and circuits
make your own graphics on existing instance shared with you, publish them and use them as needed, learn to read the feeds of other users to also use their graphics productions
create new instance:
replicator.php
navigate to glyph editor, test:
index.php
go to tree, create a new branch on it:
tree.php
go to that branch and start editing it:
index.php?path=mynewpath/
edit the shape table:
shapetableeditor.php?path=mynewpath/
shapes.txt
pastebin.com 
actions, then symbols
index.php?path=mynewpath/
(PUBLISH)
svg/svg[datecode].svg --> web, adobe creative suite, microsoft office suite, inkscape
(PNG PUBLISH)
save as: filename.png --> powerpoint, word, latex, web
web links--> svgfeed.php --> index.php?path=mynewpath/&url=mynewpath/svg/svg[datecode].svg
share and spread to other users, teaching them how to create an instance, use it, modify it and spread it themselves so that the entire system self replicates and evolves based on user needs.

@@

@@

\section{Future Work and Potential Impact}

>>>>

\title{Geometron: A Universal Language for Graphical Communication}
\author{Lafe Spietz}

@@

    The choices made here are in some sense arbitrary: 8 could just as well be 4 or 16 or even some random prime number.  But the point of this approach to language design is to choose numbers and structures which are well suited to easily being interacted with by the human mind.  A 8x8x8 cube can be divided into ``tablets'', each of which is 8X in size.

@@

\section{Workflow}

There are three levels of interaction a user can have with Geometron.  At the minimal level of interaction, one consumes graphical content as it is made availalbe, simply using the free Public Domain vector graphics that are made available for ones own work.  The next and most common level of interaction is learning how to make your own graphics, and using an existing Geometron instance to do that, creating custom graphics for whatever media you're working on.  The third level of involvement is to build up your own graphical languages, making custom tables of symbols for whatever your technical application or branding might require.  This is where the true power of Geometron will be unleashed: when people in the community start to build and share their own new graphical languages.  One might think of this as a sort of meta-Feynman diagram: a general system for diagramatically solving new types of problem in quantum theory(or indeed any field).  It is my hope that each of these work flows leads continuously into the next, with people using the graphics, learning a little of the language, then getting hooked on how much fun it is to make interesting diagrams quickly, start making their own for fun, and eventually incorporating this into their main graphical workflow, and finally when they're very comfortable in the language, building their own symbolic languages(which is the same commands you use to make any symbol.)

The graphical languages I focus on on www.quantumart.org are quantum circuits(circuits with Josephson juncitons), quantum logic gates(based on the introduction to Nielson and Chuang), Feynman diagrams, and simple knot theory for topological applications.  The static libraries just focus on the quantum circuits and the logic gates to keep the signal to noise ratio of the library high.

@@

Full Workflow:
\begin{enumerate}

\item Read this paper, learn goals of Geometron, decide to learn it.
\item Browse existing graphics libraries, download whatever vector graphics or bitmap images are of use to you for your own library.
\item Navigate your Web browser to an existing Geometron instance, try some basic operations with your keyboard to get a feel for how it works, move the cursor around, delete and edit glyphs, draw circles and lines, and some shapes
\item Make or have made your own custom Geometron keyboard for a further commitment to deplying it
\item Explore the system to make sure you can make the kinds of figures you need for your applications, learn to make and export and use PNG's and SVG's
\item Install your own instance of Geometron on your web servers or a local host inside your firewall, generally a Raspberry Pi
\item Rework the look and feel of the front page of your instance for the branding of your group/company/institution, working in collaboration with your front end Web developer team 
\item Construct your own shape libraries for the specific applications of your lab and your research and your technology and branding.  Share as much as you feel comfortable with, hopefully enlarging the graphical commons shared by the whole community.  
\item Create and share your own media feeds of SVG and png files with the rest of the world, sharing links to either your original graphics feed or a mirror of it if you're working behind a firewall.
\item Document your modifications to the code, spread the ideas to other quantum information groups, and teach them to replicate this entire workflow.

\end{enumerate}

First read this paper, focusing especially on the figures, and noting any errors or corrections to send to the author or further notes you want to add in your own documentation later.

The graphics library can be found at www.quantumart.org/library.  Peruse it at your leisure, share it widely use it as needed.

To start with movements, load the browser to the page www.quantumart.org/sandbox, and the cursor should automatically be placed in the correct input for actions to happen so that all you need to do is hit keys.  Start with the keys a,s, d, and f(all lowercase) as wel as g, h, j and k.  These should move the cursor around, and give you a chance to see how that works.  Use the backspace key to delete an action from the glyph.  Use the left and right arrow keys to navigate the cursor through the action sequence of your glyph.  When the movement makes some sense, add the basic drawing actions z,x,c and v which are point, circle, line segment and arc.

After these basic motions make sense, keep plyaing with them to try making some basic shapes and cartoons, but this will remain limited until you get used to it, and it's a good time to move on to the symmetries.  The three you mostly wan to play with are 90 degrees which is the back tic mark under the tilde in the far left of the top row of a standard ASDF keyboard, the 1 and the 2 keys, which are 72 degrees and 60 degrees respectively.  Also, 3 is bisect angle and 4 is double, 5 is trisect and 6 is triple angle.

Try differnt scale by going along the row from the angle commands to 7,8,9,0,- and = keys, which are all scaleFactor values.  Try different scale values and then zoom in and out with j and k to see how they work.

make a Bezier

make a closed path, filled and unfilled

make text with 01xxx, with 0365

choose layers, use them

edit style JSON

export a png, save it to your computer

svae a svg, link to it, save it to your computer

publish a svg link, share it so others can edit on server and publish edits and share them

Build your own space on quantumart.org/yournamehere

Build your own server, hire me, build a pi locally, build graphics, share and publish on web using a mirror, customize the language, build a physical node for the network in a conference room with a custom keyboard and a projector on a big screen

Customize your code for your brand and your research, publish it all, share your artwork as art

@@

Read this paper(or skim it and look at the pictures) 
download free graphics for use in your documents from graphics library
try out basic movements, drawing actions, make circles, lines, and gates and circuits
make your own graphics on existing instance shared with you, publish them and use them as needed, learn to read the feeds of other users to also use their graphics productions
create new instance:
replicator.php
navigate to glyph editor, test:
index.php
go to tree, create a new branch on it:
tree.php
go to that branch and start editing it:
index.php?path=mynewpath/
edit the shape table:
shapetableeditor.php?path=mynewpath/
shapes.txt
pastebin.com 
actions, then symbols
index.php?path=mynewpath/
(PUBLISH)
svg/svg[datecode].svg $\rightarrow$ web, adobe creative suite, microsoft office suite, inkscape
(PNG PUBLISH)
save as: filename.png $\rightarrow$ powerpoint, word, latex, web
web links$\rightarrow$ svgfeed.php $\rightarrow$ index.php?path=mynewpath/&url=mynewpath/svg/svg[datecode].svg
share and spread to other users, teaching them how to create an instance, use it, modify it and spread it themselves so that the entire system self replicates and evolves based on user needs.

@@

replication is covered in the next white paper.
applying the concepts of geometron to quantum information processing is covered in the third white paper, quantum modeling in HTML.

@@

\section{Future Work and Implications}

Try to imagine learning about quantum two level systems without the Bloch sphere, or basic gates without the graphical represntation of control not or quantum electrodynamics without Feynman diagrams.  These graphical tools are some of the most powerful we have at our disposal(including computers and lab equipment even) as quantum information scientists.  Imagine if the bar to build a new graphical language like this were lowered to where a graduate student could easily make a new one on their smart phone while waiting for the bus.  We have no idea just how much computational power will be unleashed by creating a free open graphical network for this kind of information, but the only way to find out is to build it.

The work presented in this paper is simply a factory for graphics.  The next of the three white papers that present this work will deal with building a social network based on sharing information created with Geometron.  It is worth noting here that the World Wide Web was originially built by Tim Bernars Lee at CERN specifically to help physicists communicate with one another and to organize their research information.  I believe that it is a rational next step for the Web to build another such extension.  This next white paper will describe the Geomtron Network, which is a self-replicating system that allows for totally free rapid sharing of graphics and integration of math-formatted text with them.  It is integrated into a slide sharing system which can replace PowerPoint for rapid online sharing of graphics, and a paper authoring system which integrates with $\LaTeX$, and a system for making vector graphics from standard function plots we use frequently.  The key idea of this system is that the current most powerful form of communication of technical ideas is simple text fields superposed on top of simple graphics, and that the faster we can make and share these with the minimum possible friction, the more powerful our communication will be.  This whole system is also predecated on the idea that the future of information technology is self replication and evolution, that is the user runs simple code that replicates the whole code base to a server, where they can easily modify the code as they go along, then when they spread it to the next user, their new, modified code base is passed along, so there is no one centralized code base, but a rhizomnal network of flowing code through the network.

The third of the white papers that makes up the sequence of Geometron for quantum information will discuss how Geometron can be used for creating sequences of discrete geometric actions in Hilbert space as well as in three dimensions just as it does in two dimensions in the system discussed here.  The thesis of this paper is that it should be possible to create Geometron instances for 3d virtual reality modeling of problems which are amenable ot quantum speedup in the browser, build another geomtron instance for the space spanned by the qubits in the quantum processor, then by building direct maps between these spaces to create a user interface for people solving these problems which truly leverages the power of Hilbert space geometry, without relying on models based on classical computer science.   I will describe a system for controlling and programming a quantum processor whereby a group of users simultaneously interact with the processor in real time using a multiplayer virtual reality interface, programming the machine using yet another geometron instance which describes qubit control pulses using discrete geometry.  Finally, by showing that the output of the quantum processor can also be geometron glyphs, I will show that this is a path to the self-programming quantum computer.
